---
title: "Gumball Themes"
author: "Alex Hallam"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package had three goals/motivations.

1. Quantile frames - for quick insights
2. Color blind friendly colors - be a decent human being, make plots everyone can interpret
3. Readable Fonts - easy reading as asthetics

This vignette will show the following uses of quantile frames.

1. Scatter plots
2. Time series plots
3. Boxplots

Though there may be other uses of quantile framed plots. These are the ones that
I have historically used.

## Scatter Plots

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r include=F}
library(grid) # still working on putting this in NAMESPACE
library(gumballthemes)
library(tidyverse)
```

```{r, fig.align="center"}
#library(grid) # still working on putting this in NAMESPACE
#library(gumballthemes)
#library(tidyverse)
mtcars %>%
  as_tibble() %>%
  ggplot(aes(y = mpg, x = hp, color = factor(carb))) +
  geom_point() +
  geom_quantileframe() +
  scale_y_quantile(mtcars$mpg) +
  scale_x_quantile(mtcars$hp, digits = 0) +
  ylab("Miles Per Gallon") +
  xlab("Horse Power") +
  guides(col = guide_legend(ncol= 6)) +
  scale_color_gumball() +
  theme_gumball()
```

The above plot uses all of the functions currently provided by `gumballthemes`:

`scale_y_quantile` 

`scale_x_quantile`

`theme_gumball`

The data-pixel ratio is high in the above plot. The range of each of the axes 
are the min and max values. The thick bar represents the 25th, 50th (aka median), 
and 75th percentiles. The axes are essentially simple boxplots.

## Time Series Plots

In this example the [`fredr`](http://sboysel.github.io/fredr/articles/fredr.html) 
packages is required along with an API key from [FRED](https://research.stlouisfed.org/).
```{r, echo=F, include=FALSE}
library(fredr)
fredr_set_key("c1856e7bf91eaca807aa6584e2614397")
```

```{r}
#library(fredr)
#fredr_set_key("abcdefghijklmnopqrstuvwxyz123456")
unrate <- fredr_series_observations(
  series_id = "UNRATE",
  observation_start = as.Date("1990-01-01"),
  observation_end = as.Date("2018-04-01"),
  frequency = "q",
  units = "chg"
)

unrate %>% glimpse()
```

After pulling data a time series plots is created. Note that `sides = "l"`. 
This is because a quantile on the time axis does not make sense in this 
example. 

```{r, fig.align="center", fig.width=5.4, fig.height=3}
unrate %>% 
  ggplot(aes(y = value, x = date)) +
  geom_line() +
  geom_quantileframe(sides = "l") +
  scale_y_quantile(unrate$value) + 
  scale_x_date(date_breaks = "5 year", date_labels = "%Y") +
  ylab("Change In UNRATE") +
  xlab("") +
  annotate("text", y = 1.0, x = as.Date("1997-01-01"),
           label = "Change in UNRATE\nis normally between\n0.07% and -.20% with a\nmedian of -0.07") +
  guides(col = guide_legend(ncol= 6)) +
  scale_color_gumball() +
  theme_gumball()
```

## Boxplots

This section needs more work -- better examples.

```{r, fig.align="center", fig.width=5.4, fig.height=3}
unrate_monthly <- fredr_series_observations(
  series_id = "UNRATE",
  observation_start = as.Date("1990-01-01"),
  observation_end = as.Date("2018-04-01"),
  frequency = "m"
)

unrate_monthly %>% 
  mutate(year = factor(lubridate::year(date))) %>% 
  ggplot(aes(y = value, x = year)) +
  ggthemes::geom_tufteboxplot() +
  geom_quantileframe(sides = "l") + 
  scale_y_quantile(unrate_monthly$value, digits = 2) + 
  theme_gumball() +
  theme(axis.text.x = element_text(hjust = 1,angle = 45))

```



