---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

```{r, message=FALSE, warning=FALSE}
pacman::p_load(tidyverse, polite, scales, ggimage, ggforce,
               understatr,
               jsonlite, xml2, qdapRegex, stringi, stringr,
               rvest, glue, extrafont, ggrepel, magick, ggtext)
loadfonts(quiet = TRUE)
```


# understat aux functions

```{r}
get_script <- function(x) {
  as.character(html_nodes(x, "script"))
}

# subset data element of html page
get_data_element <- function(x, element_name) {
  stri_unescape_unicode(str_subset(x, element_name))
}

# fix json element for parsing
fix_json <- function(x) {
  str_subset(
    unlist(
      rm_square(
        x, extract = TRUE, include.markers = TRUE
      )
    ),
    "\\[\\]", negate = TRUE
  )
}

# get player name part of html page
get_player_name <- function(x) {

  player_name <- html_nodes(x, ".header-wrapper:first-child")
  trimws(html_text(player_name))
}
```


# Data

## base data

```{r}
EPL2019 <- get_league_teams_stats("EPL", "2019")

glimpse(EPL2019)

EPL2019_teams <- EPL2019 %>% 
  select(team_name) %>% unique() %>% pull() %>% 
  str_replace_all(" ", "_")

glimpse(EPL2019_teams)
```


```{r}
get_team_shots_data <- function(team_name) {
  
  url <- glue::glue("https://understat.com/team/{team_name}/2019")
  
  team_page <- polite::bow(url)
  
  team_data <- polite::scrape(team_page) %>% 
    get_script()
  
  team_situation_data <- get_data_element(team_data, "statisticsData") %>% 
    str_replace(., "\\{\"situation\"",  "[\\{\"situation\"") %>% 
    str_replace(., "\\}'\\)", "\\}]'\\)") %>% 
    fix_json() %>% 
    fromJSON()
  
  ## stats by theme
  
  situation_df <- team_situation_data %>% 
    .[[1]] %>% 
    unlist() %>% 
    enframe()
  
  formation_df <- team_situation_data %>% 
    .[[2]] %>% 
    unlist() %>% 
    enframe()
  
  gamestate_df <- team_situation_data %>% 
    .[[3]] %>% 
    unlist() %>% 
    enframe()
  
  minutes_df <- team_situation_data %>% 
    .[[4]] %>% 
    unlist() %>% 
    enframe()
  
  zone_df <- team_situation_data %>% 
    .[[5]] %>% 
    unlist() %>% 
    enframe()
  
  speed_df <- team_situation_data %>% 
    .[[6]] %>% 
    unlist() %>% 
    enframe()
  
  shot_df <- team_situation_data %>% 
    .[[7]] %>% 
    unlist() %>% 
    enframe() 
  
  data_df <- list(situation_df, formation_df, gamestate_df, 
                  minutes_df, zone_df, speed_df, shot_df)
  
  print(paste(team_name, " done!"))
  
  return(data_df)
}
```

# EPL all teams

```{r}
EPL_shots_data_df_raw <- map(EPL2019_teams, 
                             ~get_team_shots_data(team_name = .x)) %>% 
  set_names(EPL2019_teams)
```

```{r}
saveRDS(EPL_shots_data_df_raw, 
        file = here::here("data/EPL_shots_data_df_raw.RDS"))
```

```{r}
EPL_shots_data_df_raw <- readRDS(
  file = here::here("data/EPL_shots_data_df_raw.RDS"))
```




## game state


```{r}
EPL_gamestate_data_df <- EPL_shots_data_df_raw %>% 
  ## pick out 3rd data frame which is game states
  map(3) %>% 
  bind_rows(.id = "team_name") %>% 
  mutate(name = str_replace(name, "Goal diff ", ""),
         name = str_replace(name, "against.", "against_")) %>% 
  separate(name, c("game_state", "metric"), 
           sep = "\\.", extra = "merge") %>% 
  filter(metric != "stat") %>% 
  mutate(state = case_when(
    game_state == "0" ~ "Drawing",
    game_state %in% c("> +1", "+1") ~ "Winning",
    game_state %in% c("-1", "< -1") ~ "Losing",
    TRUE ~ NA_character_
  )) %>% 
  filter(metric == "time") %>% 
  select(-game_state, -metric) %>% 
  mutate(value = as.numeric(value)) %>% 
  group_by(state, team_name) %>% 
  summarize(value = sum(value)) %>% 
  ungroup() %>% 
  group_by(team_name) %>% 
  mutate(total = sum(value)) %>% 
  ungroup() %>% 
  group_by(state, team_name) %>% 
  mutate(percentage = value / total) %>% 
  ungroup() %>% 
  arrange(team_name) %>% 
  mutate(team_name = str_replace(team_name, "_", " "),
         team_name = as_factor(team_name),
         state = as_factor(state),
         state = fct_relevel(state, "Losing", "Drawing", "Winning"),
         percentage = percentage * 100)

glimpse(EPL_gamestate_data_df)
```

```{r}
result_vals <- c("Winning" = "darkgreen", "Drawing" = "grey", "Losing" = "red")

EPL_gamestate_df_clean <- EPL_gamestate_data_df %>% 
  pivot_wider(names_from = "state", values_from = "percentage") %>% 
  mutate(Winning = if_else(is.na(Winning), 0, Winning),
         Drawing = if_else(is.na(Drawing), 0, Drawing),
         Losing = if_else(is.na(Losing), 0, Losing)) %>% 
  group_by(team_name) %>% 
  summarize_all(sum) %>% 
  ungroup() %>% 
  mutate(equals = Winning + Drawing + Losing) %>% 
  select(-equals, -value, -total) %>% 
  mutate(win_half = Winning / 2,
         draw_half =(  (Winning + Drawing) - Winning  )  / 2 + (Winning),
         lose_half = ( (Winning + Drawing + Losing) - (Winning + Drawing) )  / 2 + (Winning + Drawing)) %>% 
  mutate(team_name = fct_relevel(team_name, team_winning_order),
         team_name = fct_rev(team_name)) %>% 
  # mutate(Winning = round(Winning, digits = 1),
  #        Drawing = round(Drawing, digits = 1),
  #        Losing = round(Losing, digits = 1)) %>% 
  mutate(equals = Winning + Drawing + Losing) 

ars <- c(43.95393, 22.45681, 	33.58925)
sum(ars)

largeRem2(ars)
pcVec <- ars
```


```{r}
ggplot(EPL_gamestate_df_clean,
       aes(x = team_name, xend = team_name)) + 
  geom_segment(aes(y = 0, yend = Winning, 
                   color = "Winning"), 
               size = 4) +
  geom_segment(aes(y = Winning, yend = Winning + Drawing, 
                   color = "Drawing"), 
               size = 4) +
  geom_segment(aes(y = Winning + Drawing, yend = Winning + Drawing + Losing, 
                   color = "Losing"), 
               size = 4) +
  geom_text(aes(y = win_half, label = Winning)) +
  scale_color_manual(values = result_vals, name = "Game State",
                    breaks = c("Winning", "Drawing", "Losing"),
                    labels = c("Winning", "Drawing", "Losing")) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Percentage of Time Spent:") +
  coord_flip() +
  theme_minimal() +
  theme(text = element_text(family = "Titillium Web"),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.ticks.y = element_line(color = "black"),
        axis.ticks.length.y = unit(0.25, "cm"),
        legend.position = "top")
```






```{r}
EPL_gamestate_data_df %>% 
  pivot_wider(names_from = "state", values_from = "percentage") %>% 
  mutate(team_name = fct_reorder(team_name, Winning)) %>% 
  filter(!is.na(Winning)) %>% 
  select(-Drawing, -Losing) %>% 
  arrange(desc(Winning)) %>% 
  select(team_name) %>% pull() %>% as.character() -> team_winning_order

```


```{r}
result_vals <- c("Winning" = "darkgreen", "Drawing" = "grey", "Losing" = "red")


EPL_gamestate_data_df %>% 
  mutate(team_name = fct_relevel(team_name, team_winning_order),
         team_name = fct_rev(team_name)) %>% 
  ggplot(aes(x = team_name, y = percentage, 
             fill = state)) + 
  geom_col() +
  scale_fill_manual(values = result_vals, 
                    breaks = c("Winning", "Drawing", "Losing"),
                    labels = c("Winning", "Drawing", "Losing")) +
  coord_flip() +
  theme_minimal()
```






## by minutes (4)

```{r}
EPL_minutes_df <- EPL_shots_data_df_raw %>% 
  ## pick out 4th data frame which is minutes
  map(4) %>% 
  bind_rows(.id = "team_name")

EPL_minutes_df_clean <- EPL_minutes_df %>% 
  mutate(name = str_replace(name, "against.", "against_"),
         team_name = str_replace(team_name, "_", " "),
         team_name = as_factor(team_name)) %>% 
  separate(name, c("timing", "metric"), 
           sep = "\\.", extra = "merge") %>% 
  #select(-timing) %>% 
  filter(metric != "stat") %>% 
  pivot_wider(names_from = "metric", values_from = "value") %>% 
  mutate_at(vars(-team_name, -timing), as.numeric) %>% 
  mutate(xGperShot = xG / shots,
         xGAperShot = against_xG / against_shots)

glimpse(EPL_minutes_df_clean)
```

### shots per game

```{r}
EPL_minutes_df_clean %>% 
  mutate(shots = shots / 11,
         against_shots = against_shots / 11) %>% 
  ggplot(aes(x = timing)) +
  geom_point(aes(y = shots), color = "darkgreen") +
  geom_line(aes(y = shots, group = 1), color = "darkgreen") +
  geom_point(aes(y = against_shots), color = "red") +
  geom_line(aes(y = against_shots, group = 1), color = "red") +
  labs(x = "Minutes", y = "Shots per Game") +
  theme_minimal() +
  facet_wrap(~ team_name)
```


### xG per shot

```{r}
EPL_minutes_df_clean %>% 
  ggplot(aes(x = timing)) +
  geom_point(aes(y = xGperShot), color = "darkgreen") +
  geom_line(aes(y = xGperShot, group = 1), color = "darkgreen") +
  geom_point(aes(y = xGAperShot), color = "red") +
  geom_line(aes(y = xGAperShot, group = 1), color = "red") +
  labs(x = "Minutes", y = "Expected Goals (per shot)") +
  theme_minimal() +
  facet_wrap(~team_name)
```





```{r}
EPL_minutes_df_clean %>% 
  mutate(goalsperGame = goals / 11,
         against_goalsperGame = against_goals / 11) %>% 
  ggplot(aes(x = timing)) +
  # actual
  geom_point(aes(y = goalsperGame), color = "darkgreen") +
  geom_line(aes(y = goalsperGame, group = 1), color = "darkgreen") +
  geom_point(aes(y = against_goalsperGame), color = "darkred") +
  geom_line(aes(y = against_goalsperGame, group = 1), color = "darkred") +
  ## xG
  geom_point(aes(y = xGperShot), color = "green") +
  geom_line(aes(y = xGperShot, group = 1), color = "green") +
  geom_point(aes(y = xGAperShot), color = "red") +
  geom_line(aes(y = xGAperShot, group = 1), color = "red") +
  labs(x = "Minutes", y = "Goals per Game | xG per Shot") +
  theme_minimal() +
  facet_wrap(~team_name)
```

waaayyy tooo messsyyyyyyy

maybe just do FOR and AGAINST separately?

or do xG Goal difference?

do goal per shot?

### for

```{r}
EPL_minutes_df_clean %>% 
  mutate(goalsperGame = goals / 11,
         against_goalsperGame = against_goals / 11) %>% 
  ggplot(aes(x = timing)) +
  # actual
  geom_point(aes(y = goalsperGame), color = "darkgreen") +
  geom_line(aes(y = goalsperGame, group = 1), color = "darkgreen") +
  ## xG
  geom_point(aes(y = xGperShot), color = "green") +
  geom_line(aes(y = xGperShot, group = 1), color = "green") +
  labs(x = "Minutes", y = "Goals per Game | xG per Shot") +
  theme_minimal() +
  facet_wrap(~team_name)
```

Leicester/Chelsea/City/LFC all vastly scoring more than expected given the quality of their chances >>> elite/on-form strikers (Vardy, Tammy, Fab-3, etc.)

city pepper opponent with many shots of middling quality but with elite players they have they are bound to score a couple eventually

burnley able to get a lot of goals late on in games despite xG staying relatively stable throughout the course of the game

### against

```{r}
EPL_minutes_df_clean %>% 
    mutate(goalsperGame = goals / 11,
         against_goalsperGame = against_goals / 11) %>% 
  ggplot(aes(x = timing)) +
  # actual
  geom_point(aes(y = against_goalsperGame), color = "darkred") +
  geom_line(aes(y = against_goalsperGame, group = 1), color = "darkred") +
  ## xG
  geom_point(aes(y = xGAperShot), color = "red") +
  geom_line(aes(y = xGAperShot, group = 1), color = "red") +
  labs(x = "Minutes", y = "Goals per Game | xG per Shot") +
  theme_minimal() +
  facet_wrap(~team_name)
```

As seen vs. LFC, Villa giving up goals late in the game! Defensive efforts == exhausting and dam breaks late on?

LFC leaking goals late game despite only allowing poor quality chances...


### goal difference

```{r}
EPL_minutes_df_clean %>% 
  mutate(xGD = xG - against_xG,
         GD = goals - against_goals) %>% 
  ggplot(aes(x = timing)) +
  # actual
  geom_point(aes(y = GD), color = "black") +
  geom_line(aes(y = GD, group = 1), color = "black") +
  ## xG
  geom_point(aes(y = xGD), color = "gray") +
  geom_line(aes(y = xGD, group = 1), color = "gray") +
  geom_hline(yintercept = 0, color = "red") +
  labs(x = "Minutes", y = "Goal Difference | xGD") +
  theme_minimal() +
  facet_wrap(~team_name)
```

As seen from other viz, LFC start slow then hit their peak near halftime and slow down as the game ends


Chelsea giving up lots of late goals... Lampard's young Chelsea not closing games well + tire from unrefined pressing system?

Man. United seem to have bursts of energy, mainly peaking in the middle of either half...


as seen in previous viz, Burnley dig themselves in a hole in first half but usually manage to claw their way back near the end of the game.

Newcastle do worse over time until they wake up in the final ~15 minutes...


Everton unlucky (?) in that they give up goals with low xG late on in the game, costing them points...


goal-contribution matrix for TIME-INTERVALS?



# Liverpool





```{r}
url <- "https://understat.com/team/Liverpool/2019"

team_page <- polite::bow(url)

team_data <- polite::scrape(team_page) %>% 
  get_script()

team_situation_data <- get_data_element(team_data, "statisticsData") %>% 
  str_replace(., "\\{\"situation\"",  "[\\{\"situation\"") %>% 
  str_replace(., "\\}'\\)", "\\}]'\\)") %>% 
  fix_json() %>% 
  fromJSON()
```




## situation (1)

```{r}
situation_df <- team_situation_data %>% 
  .[[1]] %>% 
  unlist() %>% 
  enframe()

situation_df_clean <- situation_df %>% 
  mutate(name = str_replace(name, "against.", "against_")) %>% 
  separate(name, c("situation", "metric"), 
           sep = "\\.", extra = "merge")
```

```{r}
situation_df_clean %>% 
  ggplot(aes(x = situation, y = value)) +
  geom_segment(aes(xend = situation,
                   y = 0, yend = value)) +
  geom_point(color = "red", size = 2.5) + 
  theme_minimal() +
  facet_wrap("metric", scales = "free_y")
```




## formation (2)

```{r}
formation_df <- team_situation_data %>% 
  .[[2]] %>% 
  unlist() %>% 
  enframe()
```


## game state (3)

```{r}
gamestate_df <- team_situation_data %>% 
  .[3] %>% 
  unlist() %>% 
  enframe()

gamestate_df_clean <- gamestate_df %>% 
  mutate(name = str_replace(name, "gameState.Goal diff ", ""),
         name = str_replace(name, "against.", "against_")) %>% 
  separate(name, c("game_state", "metric"), 
           sep = "\\.", extra = "merge") %>% 
  filter(metric != "stat")

glimpse(gamestate_df_clean)
```

https://experimental361.com/2019/10/08/game-states-8-oct-2019/

```{r}
game_state_clean <- gamestate_df_clean %>% 
  filter(metric == "time") %>% 
  mutate(state = case_when(
    game_state == "0" ~ "Drawing",
    game_state %in% c("> +1", "+1") ~ "Winning",
    game_state %in% c("-1", ">-1") ~ "Losing",
    TRUE ~ NA_character_
  )) %>% 
  select(-game_state, -metric) %>% 
  mutate(value = as.numeric(value)) %>% 
  group_by(state) %>% 
  summarize(value = sum(value)) %>% 
  ungroup() %>% 
  mutate(team_name = "Liverpool",
         total = sum(value),
         percentage = value / total)

glimpse(game_state_clean)
```


```{r}
ggplot(game_state_clean,
       aes(x = team_name, y = percentage, fill = as.factor(state))) + 
  geom_col()
```



## by minutes (4)

```{r}
minutes_df <- team_situation_data %>% 
  .[4] %>% 
  unlist() %>% 
  enframe()

minutes_df_clean <- minutes_df %>% 
  mutate(name = str_replace(name, "against.", "against-")) %>% 
  separate(name, c("timing", "minutes", "metric"), 
           sep = "\\.", extra = "merge") %>% 
  select(-timing) %>% 
  filter(metric != "stat") %>% 
  pivot_wider(names_from = "metric", values_from = "value") %>% 
  mutate_at(vars(-minutes), as.numeric) %>% 
  mutate(xGperShot = xG / shots,
         xGAperShot = `against-xG` / `against-shots`)

glimpse(minutes_df_clean)
```

## shot zone (5)

```{r}
zone_df <- team_situation_data %>% 
  .[[5]] %>% 
  unlist() %>% 
  enframe()

zone_df_clean <- zone_df %>% 
  mutate(name = str_replace(name, "against.", "against-")) %>% 
  separate(name, c("zone", "metric"), 
           sep = "\\.", extra = "merge") %>% 
  filter(metric != "stat") %>% 
  mutate(value = as.numeric(value))
```

```{r}
zone_df_clean %>% 
  ggplot(aes(x = zone, y = value)) +
  geom_segment(aes(xend = zone,
                   y = 0, yend = value)) +
  geom_point(color = "red", size = 2.5) + 
  theme_minimal() +
  facet_wrap("metric", scales = "free_y")
```

## attack speed (6)


## shot result (7)

```{r}
shot_df <- team_situation_data %>% 
  .[[7]] %>% 
  unlist() %>% 
  enframe() 

shot_df_clean <- shot_df %>% 
  mutate(name = str_replace(name, "against.", "against-")) %>% 
  separate(name, c("result", "metric"), 
           sep = "\\.", extra = "merge") %>% 
  filter(metric != "stat") %>% 
  mutate(value = as.numeric(value))
```

```{r}
shot_df_clean %>% 
  ggplot(aes(x = result, y = value)) +
  geom_segment(aes(xend = result,
                   y = 0, yend = value)) +
  geom_point(color = "red", size = 2.5) + 
  theme_minimal() +
  facet_wrap("metric", scales = "free_y")
```




# Plots

```{r}
minutes_df_clean %>% 
  pivot_longer(-minutes, names_to = "metric", values_to = "value") %>% 
  ggplot(aes(x = minutes, y = value)) + 
  geom_point(color = "red", size = 2.5) + 
  geom_line(aes(group = metric)) +
  theme_minimal() +
  facet_wrap("metric", scales = "free_y")
```


```{r}
minutes_df_clean %>% 
  filter(metric == "shots" |
         metric == "against-shots") %>% 
  ggplot() + 
  geom_point(data = minutes_df_clean %>% 
               filter(metric == "shots"),
             aes(x = minutes, y = value),
             color = "blue", size = 2.5) + 
  geom_point(data = minutes_df_clean %>% 
               filter(metric == "against-shots"),
             aes(x = minutes, y = value),
             color = "red", size = 2.5) + 
  geom_line(aes(x = minutes, y = value, group = metric)) +
  labs(title = "Shots/Shots Against by Time") +
  theme_minimal()
```




```{r}
minutes_df_clean %>% 
  filter(metric == "xG" |
         metric == "against-xG") %>% 
  ggplot() + 
  geom_point(data = minutes_df_clean %>% 
               filter(metric == "xG"),
             aes(x = minutes, y = value),
             color = "blue", size = 2.5) + 
  geom_point(data = minutes_df_clean %>% 
               filter(metric == "against-xG"),
             aes(x = minutes, y = value),
             color = "red", size = 2.5) + 
  geom_line(aes(x = minutes, y = value, group = metric)) +
  labs(title = "xG/xGA by Time") +
  theme_minimal()
```

per shot is probably more informative


```{r}
minutes_df_clean %>% 
  mutate(shots = shots / 11,
         `against-shots` = `against-shots` / 11) %>% 
  ggplot(aes(x = minutes)) +
  geom_point(aes(y = shots), color = "darkgreen") +
  geom_line(aes(y = shots, group = 1), color = "darkgreen") +
  geom_point(aes(y = `against-shots`), color = "red") +
  geom_line(aes(y = `against-shots`, group = 1), color = "red") +
  labs(x = "Minutes", y = "Shots per Game") +
  theme_minimal()
```


```{r}
minutes_df_clean %>% 
  ggplot(aes(x = minutes)) +
  geom_point(aes(y = xGperShot), color = "darkgreen") +
  geom_line(aes(y = xGperShot, group = 1), color = "darkgreen") +
  geom_point(aes(y = xGAperShot), color = "red") +
  geom_line(aes(y = xGAperShot, group = 1), color = "red") +
  labs(x = "Minutes", y = "Expected Goals (per shot)") +
  theme_minimal()
```

LFC finish each half very strongly!

LFC shut down opponents in 2nd half

```{r}
minutes_df_clean %>% 
  mutate(goals = goals / 11,
         `against-goals` = `against-goals` / 11) %>% 
  ggplot(aes(x = minutes)) +
  # actual
  geom_point(aes(y = goals), color = "darkgreen") +
  geom_line(aes(y = goals, group = 1), color = "darkgreen") +
  geom_point(aes(y = `against-goals`), color = "darkred") +
  geom_line(aes(y = `against-goals`, group = 1), color = "darkred") +
  ## xG
  geom_point(aes(y = xGperShot), color = "green") +
  geom_line(aes(y = xGperShot, group = 1), color = "green") +
  geom_point(aes(y = xGAperShot), color = "red") +
  geom_line(aes(y = xGAperShot, group = 1), color = "red") +
  labs(x = "Minutes", y = "Goals per Game | xG per Game") +
  theme_minimal()
```

